---
title: "SuperFund Contaminated Sites"
date: July 14th 2020
output: 
    html_document:
      theme: yeti
      highlight: zenburn
      code_folding: show
      self_contained: TRUE
      df_print: paged
      css: styles.css
runtime: shiny
---

```{r setup, load myData, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("myData.RData")
```

#### Superfund Tables

The EPA has been identifying superfund (highly contaminated sites) for many years. 

Starting with reviewing the active sites as they relate to the NPL (National Properties List)

201263 is the most promising table containing location and contaminant information

Contaminant of concern data from Superfund decision documents issued in fiscal years 1982-2017. Includes sites 1) final or deleted on the National Priorities List (NPL); and 2) sites with a Superfund Alternative Approach (SAA) Agreement in place. The only sites included that are 1) not on the NPL; 2) proposed for NPL; or 3) removed from proposed NPL, are those with an SAA Agreement in place.

-----

#### Quickstart EDA

```{r, include = TRUE, warning=FALSE, message=FALSE}

# loading libraries
library(tidyverse)
library(janitor)
library(naniar)

#minor cleanup table name and col_names
active_sites <- X201263_epa_current_NPL %>% 
  clean_names()

#missing data is always problematic the naniar package is built to quickly identify issues
gg_miss_var(active_sites, show_pct = TRUE)

#check the types and structure of the data
glimpse(active_sites)

```

-----

#### Determining an appropriate key

Extracting the unique contaminants is a good start along with the unique sites for visualizing the data. 

There are no values available in the dataset regarding contamination levels and the structure is a long format with multiple contaminants being assigned to the same site.
 
```{r, include=TRUE, warning=FALSE, message=FALSE}
# tease out the unique contaminants
unique_contaminants <- 
  unique(active_sites$contaminant_name) %>% 
  tibble() 
# replace the column name that was dropped
colnames(unique_contaminants) <- c("contaminant_name")

# create a key for each location by combining the longitude and latitude columns
unique_sites <- active_sites %>% 
  mutate(site_id = paste(longitude, latitude, sep = ","))

# create a tibble with the unique id 
unique_sites <-
  unique(unique_sites$site_id) %>% 
  tibble()

# replace the column names that were dropped
colnames(unique_sites) <- c("site_id")

# split the key column in order to plot the locations on a map
unique_sites <- unique_sites %>% 
  separate(col = "site_id", into = c("lng", "lat"),sep = ",") %>% 
  mutate(lng = as.numeric(lng), lat = as.numeric(lat)) %>% 
  filter(!is.na(lng), !is.na(lat))

```

-----

#### Leaflet Plot

I am a visual person, inspiration comes through visualization. I've prepared a leaflet map that shows the locations of the superfund sites. Black points are appropriate for the time being. I will be using the leaflet below in the destined shiny application 

```{r, include=TRUE, warning=FALSE, message=FALSE, fig.align="center"}

library(leaflet)
library(leaflet.extras)

# prepare the static leaflet map
lmap <- leaflet(data = unique_sites) %>% 
  addProviderTiles(provider = "Esri.NatGeoWorldMap") %>% 
  addCircleMarkers(lng = ~lng, lat = ~lat, opacity = 0.5, color = "#000000", radius = 2) %>% 
  setView(lng = -97.909299, lat = 36.395046, zoom = 4) %>% 
  addResetMapButton() %>% 
  addSearchGoogle()
# calling the map
lmap

```

-----

#### Data Cleaning (Ongoing)

Constructing a coherent plan for classifying the data begins with identifying the information from the existing data.

Checking the boxes (What we do know)

1. Location (Y)
2. Contaminant (Y)
3. Type of Material Contaminated (Y)
4. Active Sites (Y) *on the NPL list*

What we don't know

1. Health Affects (N) *exploring EPA database for contaminants*, *pubchem JSON files*
  *How to classify health affects (Carcinogen, Mutating, Long Term)
2. Level of Contamination (N)
3. Environmental impact (N) 

-----

Stay Tuned, I'll be documenting the complete process including dead ends