---
title: "EPA National Properties Superfund Sites"
date: July 14th 2020
output: 
    html_document:
      theme: yeti
      highlight: zenburn
      code_folding: show
      self_contained: TRUE
      df_print: paged
      css: styles.css
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("myData.RData")
```

### Superfund Data

The EPA has been identifying and managing the cleanup of [Superfund](https://www.epa.gov/superfund/superfund-data-and-reports) (highly contaminated sites) for 40 years. Prior to the EPA there was no regulatory body that could control polution.

The data used for this project was embedded into the 
-----

#### Data Cleaning

```{r, include=TRUE, message=FALSE, warning=FALSE}
# required libraries
library(tidyverse)
library(lubridate)
library(janitor)
library(tidylo)
library(leaflet)
library(leaflet.extras)

# minor cleanup table name and col_names using the janitor package
active_sites <- EPA_current_NPL_201263 %>% 
  clean_names()

# selecting relevant columns, adjusting the text cases and touching up the contaminant names
active_sites_cln <- active_sites %>% 
  filter(npl_status == "Currently on the Final NPL") %>% 
  select(site_name, state, latitude, longitude,
         fiscal_year, media, contaminant_name) %>% 
  mutate(site_name = str_to_title(site_name),
         contaminant_name = str_to_title(contaminant_name),
         state = as.factor(state)) %>% 
  arrange(contaminant_name, media) %>% 
  mutate(contaminant_name = case_when(
    contaminant_name %in% c("C13-C18 Aliphatic Hydrocarbons",
                            "C19-C36 Aliphatic Hydrocarbons",
                            "C5-C8 Aliphatic Hydrocarbons",
                            "C9-C12 Aliphatic Hydrocarbons",
                            "C9-C18 Aliphatic Hydrocarbons") ~ "Chlorinated Aliphatic Hydrocarbons",
    contaminant_name %in% c("Heptachlorodibenzo[B,E][1,4]Dioxin (Hpcdd) (Mixed Isomers)",
                            "1,2,3,4,6,7,8-Heptachlorodibenzo[B,E][1,4]Dioxin (Hpcdd)",
                            "1,2,3,4,6,7,8,9-Octachlorodibenzo[B,E][1,4]Dioxin (Ocdd)",
                            "1,2,3,4,7,8-Hexachlorodibenzo[B,E][1,4]Dioxin (Hxcdd)",
                            "1,2,3,6,7,8-Hexachlorodibenzo[B,E][1,4]Dioxin (Hxcdd)",
                            "2,3,7,8-Tetrachlorodibenzo-P-Dioxin (Tcdd) Toxicity Equivalents (Teq)",
                            "2,3,7,8-Tetrachlorodibenzo-P-Dioxin (Tcdd)",
                            "Hexachlorodibenzo[B,E][1,4]Dioxin (Hxcdd) (Mixed Isomers)",
                            "1,2,3,7,8-Pentachlorodibenzo[B,E][1,4]Dioxin (Pecdd)",
                            "Pentachlorodibenzo[B,E][1,4]Dioxin (Pecdd) (Mixed Isomers)",
                            "Tetrachlorodibenzo[B,E][1,4]Dioxin (Tcdd) (Mixed Isomers)",
                            "Dioxins (Chlorinated Dibenzodioxins)",
                            "Chlorinated Dioxins And Furans",
                            "Dioxins And Dibenzofurans") ~ "Chlorobenzo Dioxin",
    contaminant_name %in% c("1-Methyl-2-Nitrobenzene",
                            "1-Methyl-3-Nitrobenzene",
                            "1-Methyl-4-Nitrobenzene") ~ "Methyl-Nitrobenzene", 
    contaminant_name %in% c("Total Extractable Petroleum Hydrocarbons (Teph)",
                            "Total Petroleum Hydrocarbon -Diesel",
                            "Total Petroleum Hydrocarbon -Gasoline",
                            "Total Petroleum Hydrocarbons (Tph)",
                            "Total Recoverable Petroleum Hydrocarbons (Trph)") ~ "Petroleum Hydrocarbons",
    contaminant_name %in% c("P,P'-Dde",
                            "P,P'-Ddd",
                            "P,P'-Ddt",
                            "Ddt And Metabolites") ~ "DDT",
    contaminant_name %in% c("C11-C22 Aromatic Hydrocarbons",
                            "C9-C10 Aromatic Hydrocarbons",
                            "Carcinogenic Polycyclic Aromatic Hydrocarbons (Cpah)",
                            "Polycyclic Aromatic Hydrocarbons",
                            "C9-C10 Aromatics",
                            "Polycyclic Aromatic Hydrocarbons (Pahs)",
                            "Polycyclic Aromatic Hydrocarbons, High Molecular Weight (Hpahs)",
                            "Polycyclic Aromatic Hydrocarbons, Low Molecular Weight (Lpahs)") ~ "Aromatic Hydrocarbons",
    contaminant_name %in% c("1,2-Dimethylbenzene (O-Xylene)",
                            "1,3-Dimethylbenzene (M-Xylene)",
                            "1,3 (Or 1,4)-Dimethylbenzene (M (Or P)-Xylene)",
                            "1,4-Dimethylbenzene (P-Xylene)") ~ "Methylbenzene Xylene",
    contaminant_name %in% c("1-Chloro-2-Methylbenzene (O-Chlorotoluene)",
                            "Chloromethylbenzene") ~ "Cloromethylbenzene",
    contaminant_name %in% c("1,2,3-Trimethylbenzene",
                            "1,2,4-Trimethylbenzene",
                            "1,3,5-Trimethylbenzene",
                            "Trimethylbenzene (Mixed Isomers)") ~ "Trimethylbenzene",
    contaminant_name %in% c("1,1-Dichloroethane",
                            "1,1,1-Trichloroethane",
                            "1,1,1,2-Tetrachloroethane",
                            "1,1,2-Trichloroethane",
                            "1,1,2,2-Tetrachloroethane",
                            "1,2-Dichloroethane",
                            "Chloroethane",
                            "Hexachloroethane",
                            "Pentachloroethane",
                            "Trichloroethane (Mixed Isomers)") ~ "Chloroethane",
    contaminant_name %in% c("1,2-Dibromo-3-Chloropropane",
                            "Dichloropropane1,2-",
                            "1,2,3-Trichloropropane",
                            "1,2-Dichloropropane",
                            "Dichloropropane (Mixed Isomers)") ~ "Chloropropane",
    contaminant_name %in% c("1,2,3,4-Tetrachlorobenzene",
                            "1,2-Dichlorobenzene",
                            "1,2,3-Trichlorobenzene",
                            "1,2,3,4-Tetrachlorobenzene",
                            "1,2,4-Trichlorobenzene",
                            "1,2,4,5-Tetrachlorobenzene",
                            "1,3-Dichlorobenzene",
                            "1,3,5-Trichlorobenzene",
                            "1,4-Dichlorobenzene",
                            "Chlorobenzene",
                            "Dichlorobenzene (Mixed Isomers)",
                            "Hexachlorobenzene",
                            "Pentachlorobenzene") ~ "Chlorobenzene",
    contaminant_name %in% c("1,2,3,4,6,7,8,9-Octachlorodibenzofuran",
                            "1,2,3,4,7,8-Hexachlorodibenzofuran (Hxcdf)",
                            "1,2,3,6,7,8-Hexachlorodibenzofuran (Hxcdf)",
                            "1,2,3,7,8-Pentachlorodibenzofuran",
                            "2,3,4,7,8-Pentachlorodibenzofuran (Pecdf)",
                            "2,3,7,8-Tetrachlorodibenzofuran",
                            "Pentachlorodibenzofuran (Pecdf)",
                            "1,3-Dioxo-1,3-Dihydro-2-Benzofuran-5-Carboxylic Acid",
                            "Tetrachlorodibenzofuran (Tcdf)",
                            "2-Benzofuran-1,3-Dione") ~ "Chlorodibenzofuran",
    contaminant_name %in% c("1,1,2,2-Tetrabromoethane",
                            "1,2-Dibromoethane") ~ "Bromoethane",
    contaminant_name %in% c("2-Amino-4,6-Dinitrotoluene",
                            "2,4-Dinitrotoluene",
                            "2,4,6-Trinitrotoluene",
                            "2,6-Dinitrotoluene",
                            "4-Amino-2,6-Dinitrotoluene",
                            "Dinitrotoluene (Mixed Isomers)",
                            "Nitrotoluene (Mixed Isomers)") ~ "Nitrotoluene",
    contaminant_name %in% c("Cresol (Mixed Isomers)",
                            "2-Methyl-4,6-Dinitrophenol (4,6-Dinitro-O-Cresol)",
                            "3-Methylphenol (Mixed Monochlorinated Isomers)",
                            "2,4-Dimethylphenol",
                            "4-Chloro-3-Methylphenol",
                            "2-Methylphenol (O-Cresol)",
                            "3-Methylphenol (M-Cresol)",
                            "4-Methylphenol (P-Cresol)") ~ "Cresol",
    contaminant_name %in% c("1,1-Dichloroethene",
                            "1,2-Dichloroethene (Cis And Trans Mixture)",
                            "Chloroethene (Vinyl Chloride)",
                            "Cis-1,2-Dichloroethene",
                            "Tetrachloroethene",
                            "Trans-1,2-Dichloroethene",
                            "Trichloroethene") ~ "Chloroethene",
    contaminant_name %in% c("(E)-1,3-Dichloro-1-Propene",                                       
                            "(Z)-1,3-Dichloro-1-Propene",
                            "1,3-Dichloropropene (Ez Mixture)",
                            "3-Chloroprop-1-Ene") ~ "Chloropropene",
    contaminant_name %in% c("1-Methylnaphthalene",
                            "2-Methylnaphthalene") ~ "Methylnaphthalene",
    contaminant_name %in% c("4-Cyano-1,2,3,4-Tetrahydro-Alpha-Methyl-1-Naphthaleneacetonitrile",
                            "4-Cyano-1,2,3,4-Tetrahydro-1-Naphthalene-Propionitrile" ) ~ "Cyanohydronaphthalene",
    contaminant_name %in% c("1,1,2-Trichloro-1,2,2-Trifluoroethane",
                            "1,1,2,2-Tetrachloro-1,2-Difluoroethane",
                            "1,2-Dichloro-1,1,2,2-Tetrafluoroethane") ~ "Chlorofloroethane",
    contaminant_name %in% c("2-Chlorophenol",
                            "2,3,5,6-Tetrachlorophenol",
                            "2,4-Dichlorophenol",
                            "2,4,5-Trichlorophenol",
                            "2,4,6-Trichlorophenol",
                            "Pentachlorophenol",
                            "Trichlorophenol (Mixed Isomers)") ~ "Chlorophenol",
    contaminant_name %in% c("(3r)-1-Azabicyclo[2.2.2]Octan-3-Yl Hydroxy(Diphenyl)Acetate",
                            "2-Hydroxy-2,2-Diphenylacetic Acid") ~ "Hydroxy Diphenylacetate",
    contaminant_name %in% c("2-Chloroaniline",
                            "3-Chloroaniline",
                            "4-Chloroaniline",
                            "4,4'-Methylenebis(2-Chloroaniline",
                            "Chlorophenyl Chloroaniline") ~ "Chloroaniline",
    contaminant_name %in% c("2-Nitroaniline",
                            "3-Nitroaniline",
                            "4-Nitroaniline",
                            "N-Methyl-N,2,4,6-Tetranitroaniline (Tetryl)") ~ "Nitroaniline",
    contaminant_name %in% c("3-(3,4-Dichlorophenyl)-1,1-Dimethylurea (Diuron)",    
                            "3-(4-Chlorophenyl)-1,1-Dimethylurea") ~ "Diuron",
    contaminant_name %in% c("2,4-Dichlorophenoxyacetic Acid",
                            "2,4,5-Trichlorophenoxyacetic Acid",
                            "(4-Chloro-2-Methylphenoxy)Acetic Acid") ~ "Chlorophenoxy Acetic Acid",
    contaminant_name %in% c("2,4-Dinitrophenol",
                            "2,4,6-Trinitrophenol",
                            "2-Nitrophenol",
                            "4-Nitrophenol") ~ "Nitrophenol",
    contaminant_name %in% c("2-(2,4-Dichlorophenoxy)Propanoic Acid",
                            "2-(2,4,5-Trichlorophenoxy)Propanoic Acid") ~ "Chlorophenoxy Propanoic Acid",
    contaminant_name %in% c("1,3-Dinitrobenzene",
                            "1,3,5-Trinitrobenzene",
                            "1,4-Dinitrobenzene",
                            "Methyl-Nitrobenzene",
                            "O-Dinitrobenzene") ~ "Nitrobenzene",
    contaminant_name %in% c("4-(4-Amino-3-Methylphenyl)-2-Methylaniline",
                            "2-Methylaniline",
                            "N,N-Dimethylaniline") ~ "Methylaniline",
    contaminant_name %in% c("4-(4-Amino-3-Chlorophenyl)-2-Chloroaniline",
                            "4,4'-Methylenebis(2-Chloroaniline)") ~ "Chloroaniline",
    contaminant_name %in% c("1-Chloro-2-[(2-Chloroethyl)Sulfanyl]Ethane",         
                            "1-Chloro-2-Ethenoxyethane") ~ "Cloro Ethenoxyethane",
    contaminant_name %in% c("(2-Methyl-2-Propanyl)Benzene",
                            "1-Methyl-4-Propan-2-Ylbenzene") ~ "Methylpropanol Benzene",
    contaminant_name %in% c("2-Methoxy-2-Methylpropane (Mtbe)",
                            "2-Methyl-2-Propanol",
                            "2-Methylprop-2-Enenitrile") ~ "Methylpropanol",
    contaminant_name %in% c("Methylcyclohexanol (Mixed Isomers)",
                            "3,5,5-Trimethylcyclohex-2-En-1-One") ~ "Methylcyclohexanol",
    contaminant_name %in% c("(2z)-2-Butenedioic Acid",
                            "4-(2,4-Dichlorophenoxy)Butanoic Acid") ~ "Butenedioic Acid",
    contaminant_name %in% c("Methyl 2-Methylprop-2-Enoate",
                            "Methyl Prop-2-Enoate") ~ "Methyl Prop Enoate",
    contaminant_name %in% c("2-Chloronaphthalene",
                            "N,N-Bis(2-Chloroethyl)-2-Naphthalenamine") ~ "Chloronaphthalene",
    str_detect(contaminant_name, "^Aroclor") ~ "PCB",
    str_detect(contaminant_name, "^Actinium") ~ "Actinium",
    str_detect(contaminant_name, "^Aluminum") ~ "Aluminum",
    str_detect(contaminant_name, "^Arsenic") ~ "Arsenic",
    str_detect(contaminant_name, "^Barium") ~ "Barium",
    str_detect(contaminant_name, "^Beryllium") ~ "Beryllium",
    str_detect(contaminant_name, "^Benzidine") ~ "Benzidine",
    str_detect(contaminant_name, "^Bromine") ~ "Bromine",
    str_detect(contaminant_name, "^Cesium") ~ "Cesium",
    str_detect(contaminant_name, "^Chromium") ~ "Chromium",
    str_detect(contaminant_name, "^Cobalt") ~ "Cobalt",
    str_detect(contaminant_name, "^Copper") ~ "Copper",
    str_detect(contaminant_name, "^Cyanide") ~ "Cyanide",
    str_detect(contaminant_name, "^Endosulfan") ~ "Endosulfan",
    str_detect(contaminant_name, "^Endrin") ~ "Endrin",
    str_detect(contaminant_name, "^Europium") ~ "Europium",
    str_detect(contaminant_name, "^Iodine") ~ "Iodine",
    str_detect(contaminant_name, "^Lead") ~ "Lead",
    str_detect(contaminant_name, "^Manganese") ~ "Manganese",
    str_detect(contaminant_name, "^Mercury") ~ "Mercury",
    str_detect(contaminant_name, "^Nickel") ~ "Nickel",
    str_detect(contaminant_name, "^Nitrate") ~ "Nitrate",
    str_detect(contaminant_name, "^Nitrophenol") ~ "Nitrophenol",
    str_detect(contaminant_name, "^Methylnaphthalene") ~ "Methylnapthalene",
    str_detect(contaminant_name, "^Phosphorus") ~ "Phosphorus",
    str_detect(contaminant_name, "^Plutonium") ~ "Plutonium", 
    str_detect(contaminant_name, "^Polonium") ~ "Polonium",
    str_detect(contaminant_name, "^Polychlorinated Biphenyls") ~ "PCB",
    str_detect(contaminant_name, "^Radium") ~ "Radium",
    str_detect(contaminant_name, "^Radon") ~ "Radon",
    str_detect(contaminant_name, "^Ruthenium") ~ "Ruthenium",
    str_detect(contaminant_name, "^Strontium") ~ "Strontium",
    str_detect(contaminant_name, "^Thallium") ~ "Thallium",
    str_detect(contaminant_name, "^Thorium") ~ "Thorium",
    str_detect(contaminant_name, "^Uranium") ~ "Uranium",
    str_detect(contaminant_name, "^Vanadium") ~"Vanadium",
    contaminant_name == "Alpha-Hexachlorocyclohexane" ~ "Lindane",
    contaminant_name == "1,2-Dihydroacenaphthylene" ~ "Acenaphthylene",
    contaminant_name == "2-Aminopyridine" ~ "Aminopyridine",
    contaminant_name == "(2z)-2-Butenedioic Acid" ~ "Butenedioic Acid",
    contaminant_name == "Benzo[A]Pyrene Equivalents (Bapeq)" ~ "Benzo[A]Pyrene",
    contaminant_name == "1,3-Benzenediol" ~ "Benzenediol",
    contaminant_name == "1,1'-Biphenyl" ~ "Biphenyl",
    contaminant_name == "Bis(2-Chloroethoxy) Methane" ~ "Bis Chloroethoxy Methane",
    contaminant_name == "Bis(2-Chloroethyl)Ether" ~ "Bis Chloroethyl Ether",
    contaminant_name == "Bis(2-Chloroisopropyl) Ether" ~ "Bis Chloroisopropyl Ether",
    contaminant_name == "Bis(2-Ethylhexyl) Adipate" ~ "Bis Ethylhexyl Adipate",
    contaminant_name == "Bis(Chloromethyl) Ether" ~ "Bis Chloromethyl Ether",
    contaminant_name == "1-Bromo-4-Phenoxybenzene" ~ "Bromophenoxy Benzene",
    contaminant_name == "Tris(2,3-Dibromopropyl) Phosphate" ~ "Bromopropyl Phosphate",
    contaminant_name == "1-Butanol (N-Butanol)" ~ "Butanol",
    contaminant_name == "1,3-Butadiene" ~ "Butadiene", 
    contaminant_name == "1-Butoxybutane" ~ "Butoxybutane",
    contaminant_name == "2-Butoxyethanol" ~ "Butoxyethanol",
    contaminant_name == "2-Butanone (Methyl Ethyl Ketone)" ~ "Butanone Methyl",
    contaminant_name == "1,3-Butadiene" ~ "1,3-Butadiene",
    contaminant_name == "Butan-2-Ylbenzene" ~ "Diethylbenzene",
    contaminant_name == "1(2z)-2-Butenedioic Acid" ~ "Butenedioic Acid",
    contaminant_name == "9h-Carbazole" ~ "Carbazole",
    contaminant_name == "Carbonyl Dichloride (Phosgene)" ~ "Carbonyl Dichloride Phosgene",
    contaminant_name == "1,4-Benzenedicarboxylic Acid" ~ "Benzenedicarboxylic Acid",
    contaminant_name == "Chlorine (Cl2)" ~ "Chlorine",
    contaminant_name == "2,2-Dichloroethenyl Dimethyl Phosphate" ~ "Dichloroethenyl Dimethyl Phosphate", 
    contaminant_name == "1-Chloro-4-Phenoxybenzene" ~ "Chloro Phenoxybenzene",
    contaminant_name == "2,2,2-Trichloro-1,1-Bis(4-Chlorophenyl)Ethanol" ~ "Chlorophenyl Ethanol",
    contaminant_name == "4-Methylheptyl 2-(2,4,5-Trichlorophenoxy)Propanoate" ~ "Chlorophenoxy Propanoate",
    contaminant_name == "1-Bromo-4-Phenoxybenzene" ~ "Bromo Phenoxybenzene",
    contaminant_name == "10-Chloro-5h-Phenarsazinine" ~ "Chloro Phenarsazinine",
    contaminant_name == "C.i. Acid Green 3" ~ "Acid Green Colorant",
    contaminant_name == "C.i. Basic Violet 1" ~ "Basic Violet Colorant",
    contaminant_name == "Dibenzo(A,H)Anthracene" ~ "Dibenzo[AH]Athracene",
    contaminant_name == "Dibenz[A,H]Acridine" ~ "Dibenz[AH]Acridine",
    contaminant_name == "Dibenzo[A,E]Pyrene" ~ "Dibenzo[AE]Pyrene",
    contaminant_name == "Dibenzo[A,H]Pyrene" ~ "Dibenzo[AH]Pyrene",
    contaminant_name == "N,N-Dibutylnitrous Amide" ~ "Dibutylnitrous Amide",
    contaminant_name == "N,N-Diethylnitrous Amide" ~ "Diethylnitrous Amide",
    contaminant_name == "N,N-Diphenylnitrous Amide" ~ "Diphenylnitrous Amide",
    contaminant_name == "N,N-Dipropylnitrous Amide" ~ "Dipropylnitrous Amide",
    contaminant_name == "Dichloro-[(E)-2-Chloroethenyl]Arsane (Lewisite)" ~ "Lewisite",
    contaminant_name == "Diethylbenzene (Mixed Isomers)" ~ "Diethylbenzene",
    contaminant_name == "Dichloromethane (Methylene Chloride)" ~ "Methylene Chloride",
    contaminant_name == "2-(1-Methylpropyl)-4,6-Dinitrophenol (Dinoseb)" ~ "Dinoseb",
    contaminant_name == "Di-N-Octyl Phthalate" ~ "Dioctyl Phthalate",
    contaminant_name == "1,4-Dioxane" ~ "Dioxane",
    contaminant_name == "Dimethyl Phenol (Mixed Isomers)" ~ "Dimethyl Phenol", 
    contaminant_name == "Diaminotoluene (Mixed Isomers)" ~ "Diaminotoluene",
    contaminant_name == "1,2-Diphenylhydrazine" ~ "Diphenylhydrazine",
    contaminant_name == "Total Dissolved Solids" ~ "Dissolved Solids",
    contaminant_name == "1,4-Dithiane" ~ "Dithiane", 
    contaminant_name == "Ethane-1,2-Diamine" ~ "Ethane Diamine",
    contaminant_name == "2-Ethoxyethanol" ~ "Ethoxyethanol",
    contaminant_name == "Bis(2-Ethylhexyl)Phthalate" ~ "Ethylhexyl Phthalate",
    contaminant_name == "O-Ethyl O-(4-Nitrophenyl) Phenylphosphonothioate" ~ "Ethyl Nitrophenyl Phosphonothioate",
    contaminant_name == "S-Ethyl N,N-Dipropylcarbamothioate (Eptc)" ~ "Ethyl Dipropylcarbamothioate",
    contaminant_name == "Ethyl Prop-2-Enoate" ~ "Ethyl Prop Enoate",
    contaminant_name == "1,2-Ethanediol (Ethylene Glycol)" ~ "Ethylene Glycol",
    contaminant_name == "9h-Fluorene" ~ "Fluorene",
    contaminant_name == "2-Fluoroacetic Acid" ~ "Fluoroacetic Acid",
    contaminant_name == "Fluorine (F2)" ~ "Fluorine",
    contaminant_name == "2-[Fluoro(Methyl)Phosphoryl]Oxypropane (Sarin)" ~ "Sarin",
    contaminant_name == "Gamma-Chlordane" ~ "Gamma Chlordane",
    contaminant_name == "Gamma-Hexachlorocyclohexane (Lindane)" ~ "Lindane",
    contaminant_name == "Total Trihalomethanes" ~ "Halomethane",
    contaminant_name == "Hexahydro-1,3,5-Trinitro-1,3,5-Triazine (Rdx)" ~ "Cyclonite",
    contaminant_name == "Hydrogen (H2)" ~ "Hydrogen",
    contaminant_name == "2-Hexanone" ~ "Hexanone", 
    contaminant_name == "1h-Indene" ~ "Indene",
    contaminant_name == "Indeno(1,2,3-Cd)Pyrene" ~ "Indeno Pyrene",
    contaminant_name == "3,6-Dichloro-2-Methoxybenzoic Acid" ~ "Chloro Methoxybenzoic Acid",
    contaminant_name == "4-Methylchrysene" ~ "Methylchrysene",
    contaminant_name == "2-Methyloxirane" ~ "Methyloxirane",
    contaminant_name == "2,2,4-Trimethylpentane" ~ "Methylpentane",
    contaminant_name == "4-Methyl-2-Pentanone (Methyl Isobutyl Ketone)" ~ "Methylpentanone",
    contaminant_name == "4,4'-Methylenebis(2-Chloroaniline)" ~ "Methylenebis Chloroaniline",
    contaminant_name == "4-Methoxyphenol" ~ "Methoxyphenol",
    contaminant_name == "2-Naphthalenamine" ~ "Naphthalenamine",
    contaminant_name == "1-Nitrosopyrrolidine" ~ "Nitrosopyrrolidine",
    contaminant_name == "4-Nitrosodiphenylamine" ~ "Nitrosodiphenylamine",
    contaminant_name == "N-Nitrosodimethylamine" ~ "Nitrosodimethylamine",
    contaminant_name == "2,2',2''-Nitrilotriethanol" ~ "Nitrilotriethanol",
    contaminant_name == "O-Ethyl S,S-Dipropyl Phosphorodithioate (Ethoprop)" ~ "Ethoprop",
    contaminant_name == "O,O,O,O-Tetraethyl Dithiodiphosphate" ~ "Sulfotep",
    contaminant_name == "Oil & Grease" ~ "Oil and Grease",
    contaminant_name == "2,2'-Oxydiethanol" ~ "Oxydiethanol", 
    contaminant_name == "2-Propan-2-Yloxypropane" ~ "Oxypropane",
    contaminant_name == "Organic & Inorganic Liquid Sludge" ~ "Liquid Sludge",
    contaminant_name == "P-Cymene" ~ "P Cymene",
    contaminant_name == "Pentaerythritol Tetranitrate (Petn)" ~ "Pentaerythritol Tetranitrate",
    contaminant_name == "Perfluorooctanoic Acid (Pfoa)" ~ "Perfluorooctane Sulfonic Acid",
    contaminant_name == "10h-Phenothiazine" ~ "Phenothiazine",
    contaminant_name == "Polybrominated Biphenyls (Firemaster Ff 1)" ~ "Polybrominated Biphenyls",
    contaminant_name == "1-Phenylethanone" ~ "Phenylethanone",
    contaminant_name == "2-Chloro-1-Phenylethanone" ~ "Chloro Phenylethanone", 
    contaminant_name == "4-Phenylaniline" ~ "Phenylaniline",
    contaminant_name == "Promethium-147" ~"Promethium",
    contaminant_name == "2-Propanol" ~ "Propanol",
    contaminant_name == "1,2-Propanediol" ~ "Propanediol", 
    contaminant_name == "[(E)-Prop-1-Enyl]Benzene" ~ "Propenyl Benzene",
    contaminant_name == "1-Propene" ~ "Propene",
    contaminant_name == "2-Propenenitrile (Acrylonitrile)" ~ "Acrylonitrile",
    contaminant_name == "Residual Range Organics (Rro)" ~ "Residual Range Organics",
    contaminant_name == "Semi-Volatile Organic Compounds (Svoc)" ~ "Semi Volatile Organic Compounds",
    contaminant_name == "Silicon Dioxide (Amorphous Silica)" ~ "Silicon Dioxide",
    contaminant_name == "Sodium Nitrite" ~ "Sodium Nitrate",
    contaminant_name == "Sodium-22" ~ "Sodium",
    contaminant_name == "1,3,5,7-Tetranitro-1,3,5,7-Tetrazocane (Hmx)" ~ "Octogen",
    contaminant_name == "Total Suspended Solids (Tss)" ~ "Suspended Solids",
    contaminant_name == "Toluene Diisocyanate (Mixed Isomers)" ~ "Toluene Diisocyanate",
    contaminant_name == "Total Benzofluoranthenes" ~ "Benzofluoranthene",
    contaminant_name == "Total Dissolved Solids (Tds)" ~ "Total Dissolved Solids",
    contaminant_name == "Unexploded Ordnance (Uxo)" ~ "Unexploded Ordnance",
    contaminant_name == "Unknown Liq Waste" ~ "Liquid Waste Unknown",
    contaminant_name == "Voc" ~ "Volatile Organic Compounds",
    contaminant_name == "Vx" ~ "VX Nerve Agent",
    contaminant_name == "Xylene (Mixed Isomers)" ~ "Xylene",
    TRUE ~ contaminant_name
    ))

```

```{r}
# just quickly reviewing the counts per state, may need this table and will delete it if not
unique_sites_tbl <- active_sites_cln %>% 
  tabyl(contaminant_name, state)

unique_sites_tbl <- tibble(unique_sites_tbl)

```

### Data Wrangling

```{r, include=TRUE, message=FALSE, warning=FALSE}

# counting the unique contaminants per state
unique_contaminants_state <- active_sites_cln %>% 
  group_by(state) %>% 
  count(contaminant_name, sort = TRUE, name = "n.contaminants_state") %>% 
  mutate(contaminant_name = str_to_title(contaminant_name)) %>% 
  arrange(contaminant_name) %>% 
  mutate(temp_key = paste(state, contaminant_name, sep = ":"))

# counting and isolating the unique contaminants
unique_contaminants <- active_sites_cln %>% 
  count(contaminant_name, sort = TRUE, name = "n.contaminants") %>% 
  mutate(contaminant_name = str_to_title(contaminant_name)) %>% 
  bind_cols(pal_con) # colours added for the leaflet plots

# counting and isolating the unique sites
unique_sites <- active_sites_cln %>% 
  count(site_name, sort = TRUE, name = "n.sites") %>% 
  mutate(site_name = str_to_title(site_name)) %>% 
  bind_cols(pal) # colours added for the leaflet plots

```

### Joining Counts

```{r, include=TRUE, message=FALSE, warning=FALSE}

# joining up the counts to the overall dataset
active_sites_cln <- active_sites %>% 
  mutate(fiscal_year = as.numeric(fiscal_year),
         contaminant_name = str_to_title(contaminant_name),
         site_name = str_to_title(site_name)) %>% 
  left_join(unique_contaminants, by = "contaminant_name") %>% 
  left_join(unique_sites, by = "site_name", suffix = c(".contaminants", ".sites")) %>% 
  arrange(desc(n.contaminants))

```

```{r, include=TRUE, message=FALSE, warning=FALSE}

# creating a key for the purposes of joining the unique_contaminants_state
lonlat_state <- active_sites_cln %>% 
  select(state, contaminant_name, longitude, latitude, site_name, fiscal_year, media, 
         colour.contaminants) %>% 
  mutate(temp_key = paste(state, contaminant_name, sep = ":")) %>% 
  distinct_all()

# using a left join to combine the data
state_contaminants_temp <- unique_contaminants_state %>% 
  right_join(lonlat_state, by = "temp_key") %>% 
  filter(!is.na(longitude)) %>% 
  distinct_all()
# update the column names
colnames(state_contaminants_temp) <- c("state", "contaminant_name", "n.contaminants_state",
                                  "temp_key", "state_long", "contaminant", "longitude", "latitude",
                                  "site_name", "fiscal_year", "media", "colour")

```

Without any actual statistics in the table, in the shortterm I am using the tidylo package to generate weighted log odds


```{r}

# 

state_contaminants_fct <- state_contaminants_temp %>% 
  filter(!is.na(contaminant_name)) %>% 
  mutate(
    state = fct_inorder(state),
    contaminant_name = as_factor(contaminant_name)) %>% 
  count(state, contaminant_name) %>% 
  mutate(contaminant_name = tidytext::reorder_within(contaminant_name, n, state))

statelo <- state_contaminants_fct %>% 
  bind_log_odds(state, contaminant_name, n) %>% 
  arrange(-log_odds_weighted) 

unique_states <- unique(as.character(statelo$state), sort = TRUE) 

unique_states <- tibble(unique_states)

statelo_app <- statelo %>% 
  mutate(contaminant_name = as.character(contaminant_name),
         temp_key = as.character(contaminant_name),
         state = as.character(state)) %>% 
  separate(contaminant_name, sep = "___", into = c("contaminant", "state_abr"))

statelo_app <- statelo_app %>% 
  mutate(state = case_when(
    state == "AK" ~ "Alaska",
    state == "AL" ~ "Alabama",
    state == "AR" ~ "Arkansas",
    state == "AS" ~ "Samoa",
    state == "AZ" ~ "Arizona",
    state == "CA" ~ "California",
    state == "CO" ~ "Colorado",
    state == "CT" ~ "Connecticut",
    state == "DC" ~ "District of Columbia",
    state == "DE" ~ "Delaware",
    state == "FL" ~ "Florida",
    state == "GA" ~ "Georgia",
    state == "GU" ~ "Guam",
    state == "HI" ~ "Hawaii",
    state == "IA" ~ "Iowa",
    state == "ID" ~ "Idaho",
    state == "IL" ~ "Illinois",
    state == "IN" ~ "Indiana",
    state == "KS" ~ "Kansas",
    state == "KY" ~ "Kentucky",
    state == "LA" ~ "Louisiana",
    state == "MA" ~ "Massachusetts",
    state == "MD" ~ "Maryland",
    state == "ME" ~ "Maine",
    state == "MI" ~ "Michigan",
    state == "MN" ~ "Minnesota",
    state == "MO" ~ "Missouri",
    state == "MS" ~ "Mississippi",
    state == "MT" ~ "Montana",
    state == "NC" ~ "North Carolina",
    state == "ND" ~ "North Dakota",
    state == "NE" ~ "Nebraska",
    state == "NH" ~ "New Hampshire",
    state == "NJ" ~ "New Jersey",
    state == "NM" ~ "New Mexico",
    state == "NV" ~ "Nevada",
    state == "NY" ~ "New York",
    state == "OH" ~ "Ohio",
    state == "OK" ~ "Oklahoma",
    state == "OR" ~ "Oregon",
    state == "PA" ~ "Pennsylvania",
    state == "PR" ~ "Puerto Rico",
    state == "RI" ~ "Rhode Island",
    state == "SC" ~ "South Carolina",
    state == "SD" ~ "South Dakota",
    state == "TN" ~ "Tennessee",
    state == "TX" ~ "Texas",
    state == "UT" ~ "Utah",
    state == "VA" ~ "Virginia",
    state == "VI" ~ "Virginia",
    state == "VT" ~ "Vermont",
    state == "WA" ~ "Washington",
    state == "WI" ~ "Wisconsin",
    state == "WV" ~ "West Virginia",
    state == "WY" ~ "Wyoming",
    TRUE ~ state
  ))  
```

```{r}

# see what we can do

statelo_join <- statelo_app %>% 
  mutate(temp_key = paste(state_abr, contaminant, sep = ":"))

state_contaminants <- state_contaminants_temp %>% 
  left_join(statelo_join, by = "temp_key") %>% 
  select(contaminant_name, n.contaminants_state, state.y, state_abr, 
         longitude, latitude, site_name, media, n, log_odds_weighted, fiscal_year, colour) %>% 
  filter(!is.na(contaminant_name)) %>% 
  mutate(log_odds_weighted = replace_na(log_odds_weighted, 1),
         log_odds_weighted = round(log_odds_weighted, 2)) %>%
  mutate(media = case_when(
    media == "Groundwater" ~ "Ground Water",
    media == "Buildings/Structures" ~ "Buildings Structures",
    TRUE ~ media)) %>% 
  mutate(pubclink = paste0("<b><a href = 'https://pubchem.ncbi.nlm.nih.gov/#query=", contaminant_name,"' target = \"_blank\">"),
         sitelink = paste0("<b><a href = 'https://search.epa.gov/epasearch/?querytext=", site_name,
                           "&areaname=&areacontacts=&areasearchurl=&typeofsearch=epa&result_template=2col.ftl#/' target = \"_blank\">"))

colnames(state_contaminants) <- c("contaminant_name", "n.contaminants_state", "state_name",
                                  "state_abr", "longitude", "latitude", "site_name", "media",              
                                  "n.log_counts", "log_odds_weighted", "fiscal_year", "colour", 
                                  "pubclink", "sitelink")


```

-----

#### Data Cleaning (Ongoing)

Constructing a coherent plan for classifying the data begins with identifying the information from the existing data.

Checking the boxes (What we do know)

1. Location (Y)
2. Contaminant (Y)
3. Type of Material Contaminated (Y)
4. Active Sites (Y) *on the NPL list*

What we don't know

1. Health Affects (N) *exploring EPA database for contaminants*, *pubchem JSON files*
  *How to classify health affects (Carcinogen, Mutating, Long Term)
2. Level of Contamination (N)
3. Environmental impact (N) 

-----

Stay Tuned, I'll be documenting the complete process including dead ends

